---
title: "Nearby roadways"
author: "Danielle Medgyesi"
date: "12/20/2022"
output: pdf_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(ggplot2)
library(knitr)
library(data.table)
library(Hmisc)
library(ggmap)
library(sp)
library(sf)
library(geosphere)
library(googleway)
library(trajr)
#OpenStreetMap data 
library(osmdata)

options(dplyr.summarise.inform=FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message  = FALSE)


`%!in%` = Negate(`%in%`)
options(scipen = 999)

```

# OpenStreetMap

Download all street types  from osm for study region 

https://wiki.openstreetmap.org/wiki/Key:highway#Roads

- **trunk**=The most important roads in a country's system that aren't motorways. (Need not necessarily be a divided highway.)

- **primary**= The next most important roads in a country's system. (Often link larger towns.)

- **secondary**= The next most important roads in a country's system. (Often link towns.)

- **tertiary**= The next most important roads in a country's system. (Often link smaller towns and villages)

- **unclassified**= The least important through roads in a country's system – i.e. minor roads of a lower classification than tertiary, but which serve a purpose other than access to properties. (Often link villages and hamlets.) The word 'unclassified' is a historical artefact of the UK road system and does not mean that the classification is unknown; you can use highway=road for that.

- **residential**= Roads which serve as an access to housing, without function of connecting settlements. Often lined with housing.

- **service**= For access roads to, or within an industrial estate, camp site, business park, car park, alleys, etc. Can be used in conjunction with service=* to indicate the type of usage and with access=* to indicate who can use it and in what circumstances.

- **track**= Roads for mostly agricultural or forestry uses. To describe the quality of a track, see tracktype=*. Note: Although tracks are often rough with unpaved surfaces, this tag is not describing the quality of a road but its use. Consequently, if you want to tag a general use road, use one of the general highway values instead of track.

- **road**= A road/way/street/motorway/etc. of unknown type. It can stand for anything ranging from a footpath to a motorway. This tag should only be used temporarily until the road/way/etc. has been properly surveyed. If you do know the road type, do not use this value, instead use one of the more specific highway=* values.
footway= For designated footpaths; i.e., mainly/exclusively for pedestrians. This includes walking tracks and gravel paths. If bicycles are allowed as well, you can indicate this by adding a bicycle=yes tag. Should not be used for paths where the primary or intended usage is unknown. Use highway=pedestrian for pedestrianised roads in shopping or residential areas and highway=track if it is usable by agricultural or similar vehicles.

```{r}

#search study region 


boundaries <- opq(bbox = 'GH-AF') %>%
  add_osm_feature(key = 'admin_level', value=6) %>% 
  osmdata_sf %>% unique_osmdata

districts<-boundaries$osm_multipolygons

D1<- districts %>% filter(name=="Asutifi South District")

D1<- D1 %>% st_transform(4326)

#first filter buildings inside xy limits
D_box<- st_bbox(D1)

location <- D_box %>% opq()

#quesry all types of roads

all_types<- available_tags("highway")


 Streets <- location %>%
   add_osm_feature(key = "highway", 
                   value = all_types) %>%
   osmdata_sf()


Street_lines<- Streets$osm_lines



kable(Street_lines %>% data.frame()%>% group_by(highway) %>%
        summarise(n()), 
      caption = "Number of streets in study region")


ggplot()+
  geom_sf(data=Street_lines)



```


# Simulated GPS trajectory 

```{r, fig.height=5, fig.width=5}

# Create a random trajectory

set.seed(333)
trj <- TrajGenerate(n = 2880, stepLength = 2, angularErrorSd = 0.05, random = TRUE)
#trj <- TrajRotate(trj, pi / 1.5, relative = FALSE)
trj<- TrajReverse(trj)
trj<- TrajTranslate(trj, 1000, 2500)

#for illustration, use a community centroid as the home coordinates 
Home_lon<- -2.43
Home_lat<- 6.85
 
trj<- trj %>% mutate(x=Home_lon + (x / 6378000) * (200 / pi) / cos(Home_lon * pi/200), 
                     y= Home_lat  + (y / 6378000) * (200 / pi))

trj_sf <- st_as_sf(trj, coords = c("x", "y"), 
                crs = 4326)


p<- ggplot()+geom_sf(data=trj_sf)



y_range<- ggplot_build(p)$layout$panel_params[[1]]$y_range+c(-0.003,0.003)
x_range<- ggplot_build(p)$layout$panel_params[[1]]$x_range+c(-0.003,0.003)


filt_bbox <- sf::st_bbox(c(xmin = x_range[1], 
                           ymin = y_range[1], 
                           xmax = x_range[2], 
                           ymax = y_range[2]), 
                         crs = st_crs(4326))%>% 
  sf::st_as_sfc(.)



Roadfiltr<-sf::st_intersection(Street_lines, filt_bbox)



ggplot()+
  geom_sf(data=trj_sf, size=0.5)+
  geom_sf(data = Roadfiltr, aes(color=highway))+
  theme_void()+
  labs(color="Road type")


```



**In order to access satellite data from Google, users must obtain a valid Google Maps API key** 

To do so, you can create an account with Google: https://mapsplatform.google.com/

- Begin a new project and create credentials

- I created an API key, selecting "restrict key" > "Maps JavaScript API"

- API can then be called into R using the "register_google" function

- Make sure your API is secure; charges may be apply to high-volume users 

*If you do not wish to create an API, skip code below and refer to plain ggplot maps*

```{r, fig.height=6, fig.width=6}

#map_key <- ""
#register_google(key=map_key)

base<- get_map(location = c(-2.41, 6.89), zoom=14, maptype = "satellite")


ggmap(base)+
  coord_sf(crs = st_crs(4326)) +
  geom_sf(data = Roadfiltr, aes(color=highway), size=1.1,
          inherit.aes = FALSE)+
   geom_sf(data = trj_sf,color="red", size=0.5,  
          inherit.aes = FALSE)+
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.margin = unit(c(0, 0, -1, -1), 'lines')) +
  xlab('') +
  ylab('')+
  labs(color="Road type")


```


# Identify nearby roadways and density (100m buffer)

```{r}

#Transform to Ghana crs 
#https://epsg.io/?q=Ghana
Streets_sf<- Roadfiltr %>% st_transform(32630)
trj_sf <- trj_sf %>% st_transform(32630)
#create unique ID for each point
trj_sf<- trj_sf %>% mutate(ID=1:nrow(.))

#Split street dataset into list by street type 
Streets_sf_list<- split(Streets_sf, f= Streets_sf$highway)

#Create 100m buffer around GPS points
trj_sf100 <- st_buffer(trj_sf, dist = 100)


#loop through each road type
trj_roadways<- list()
for (i in 1:length(Streets_sf_list)) {
# identify roads that are within 100m
intrsct = st_intersection(trj_sf100, Streets_sf_list[[i]])
# save the length of each road segment
intrsct$len = st_length(intrsct)
# sume road length for each point
sum_len = intrsct %>% group_by(ID) %>% summarise(len = sum(len))
colnames(sum_len)[2]<- paste0(names(Streets_sf_list)[[i]], "_len")
trj_roadways[[i]]<- st_drop_geometry(sum_len)

}


#includes only observations near roadways
trj_roadways_df<- trj_roadways %>% reduce(full_join, by= "ID")

#merge with full trajectory 
trj_sf<- trj_sf %>% left_join(trj_roadways_df, by="ID")

#replace NA with 0 (e.g., roadway not near)
trj_sf<- trj_sf%>%
  mutate(across(residential_len:unclassified_len, ~ ifelse(is.na(.), 0, .)))

#create binary indicator (yes/no) roadway is within 100m 
trj_sf<- trj_sf%>% 
  mutate(across(residential_len:unclassified_len, ~ ifelse(.>0, 1, 0), 
                                 .names = "{col}BI"))


#create median split among observations near road 
trj_sf<- trj_sf%>% 
  mutate(across(c(residential_len:unclassified_len), 
          ~ cut2(., cuts = c(0, min(.[.>0]),
            median(.[.>0]), 
            max(.[.>0]))), .names = "{col}MED"))


```


# Map whether near roadway 

```{r, fig.height=5, fig.width=5}

ggplot()+
  geom_sf(data=trj_sf, aes(color=as.character(trunk_lenBI)))+
  geom_sf(data = Roadfiltr[Roadfiltr$highway=="trunk",])+
  theme_void()+
  scale_color_manual(name="Near trunk road (100m)",
                     labels=c("no", "yes"),
                     values = c("blue", "red"))



ggplot()+
  geom_sf(data=trj_sf, aes(color=as.character(residential_lenBI)))+
  geom_sf(data = Roadfiltr[Roadfiltr$highway=="residential",])+
  theme_void()+
  scale_color_manual(name="Near residential road (100m)",
                     labels=c("no", "yes"),
                     values = c("blue", "red"))



ggplot()+
  geom_sf(data=trj_sf, aes(color=as.character(unclassified_lenBI)))+
  geom_sf(data = Roadfiltr[Roadfiltr$highway=="unclassified",])+
  theme_void()+
  scale_color_manual(name="Near unclassified road (100m)",
                     labels=c("no", "yes"),
                     values = c("blue", "red"))



```


# Map by roadway density 

*< or >= median among observations near road*

```{r, fig.height=5, fig.width=5}



ggplot()+
  geom_sf(data=trj_sf, aes(color=as.character(trunk_lenMED)))+
  geom_sf(data = Roadfiltr[Roadfiltr$highway=="trunk",])+
  theme_void()+
  scale_color_manual(name="Trunk density",
                     values = c("blue","orange", "red"))




```



